---
title: "main"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
The dictionary meaning of "Democracy" defined by Encyclopedia Britannica is "literally ruled by the people." As we all learned from social science classes from junior high school, democracy is derived from the Greek words "coined from demos -people- and kratos -rule- in the middle of the 5th century BC to denote the political systems" in Athens. In modern usage, democracy often refers to a system of government where the citizens exercise power directly or through electing representatives in order to form a parliament. Thus, it is to be distinguished from monarchy or dictatorship. Yet, there is no consensus on how to define democracy while some of characteristics such as political freedom, rule of law, and equality before the law are commonly seen and identified as basic features of democracy. According to "The Economist Intelligence Unit's index of democracy" written by Laza Kekic, "the fundamental features of a democracy include government based on majority rule and the consent of the governed, the existence of free and fair elections, the protection of minorities and respect for basic human rights" (1). Also, the United Nations sees democracy as "one of the universal and indivisible core values [which] is based on the freely expressed will of people and closely linked to the rule of law and exercise of human rights and fundamental freedoms." Sometimes, the terms "freedom" and "democracy" are used interchangeably. But democracy, as a set of practices and principles about freedom, rather protects freedom and equality.   
Measuring democracy also has been contested that there still have ongoing debates on this subject. The most widely-known measure is from the US-based Freedom House organization, which aggregates scores of 25 indicators up to 100, on a 1 to 7 rating scale, from political rights and civil liberties. For the report of 2016, Freedom House evaluated the state of freedom in 195 countries and 15 territories during calendar year 2015, and 125 countries are classified as electoral democracies. The Freedom House's electoral democracy measure is a thin or minimalist whose measure of democracy may differ from think or wider measures which include various aspects of social and political culture in democratic societies. Another organization where also provides measures of democracy and regime types based on minimalist definitions is The Polity Project. Specifically focused on the institutionalized authority patterns of the state regime, The Polity Project "examines concomitant qualities of democratic and autocratic authority in governing institutions [.to] envision a spectrum of governing authority that spans from fully institutionalized autocracies through mixed, or incoherent, authority regimes to fully institutionalized democracies (Center for Systemic Peace). For our current research, this Polity data from the Polity Project will be used. The detailed information with regard to the "Polity Score" will be discussed at the "Data, Variables, and Methods" section. 
Following the introduction, this research paper will proceed by presenting background researches about democratization and its history in Africa to set up the baseline and link to our research question and hypotheses at the following section. After that, details about data, variable sections, and overall statistical methods will be addressed along with explanations how the authors collected, cleaned, and utilized each variable. Based on the cleaned data, the authors will provide descriptive analyses as a preliminary stage, and then the multivariate analyses conducted based on the various statistical models will be interpreted by concentrating on the effects of five socio-economic variables on democratization in three selected African countries. 

  

# Research Question
test


# Data Availability

1) Measurement of Democratization
  1-1) Polity 4
 ?? ??http://www.systemicpeace.org/inscr/p4manualv2015.pdf
  1-2) 

2) Gross National Income Level

3) primary enrollment

4) Income Inequality

5) Gender Inequality in labor force

6) mortality rate under 5

#Data Cleaning

#Data Analysis

# Discussion

# Conclusion
```{r}


```

