---
title: "main"
output: pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(stargazer)
library(bibtex)
library(plm)
library(knitr)
library(lmtest)
library(foreign)
library(car)
```

# Introduction
     The dictionary meaning of "Democracy" defined by Encyclopedia Britannica is "literally ruled by the people." As we all learned from social science classes from junior high school, democracy is derived from the Greek words "coined from demos -people- and kratos -rule- in the middle of the 5th century BC to denote the political systems" in Athens. In modern usage, democracy often refers to a system of government where the citizens exercise power directly or through electing representatives in order to form a parliament. Thus, it is to be distinguished from monarchy or dictatorship. Yet, there is no consensus on how to define democracy while some of characteristics such as political freedom, rule of law, and equality before the law are commonly seen and identified as basic features of democracy. According to "The Economist Intelligence Unit's index of democracy" written by Laza Kekic, "the fundamental features of a democracy include government based on majority rule and the consent of the governed, the existence of free and fair elections, the protection of minorities and respect for basic human rights" (1). Also, the United Nations sees democracy as "one of the universal and indivisible core values [which] is based on the freely expressed will of people and closely linked to the rule of law and exercise of human rights and fundamental freedoms." Sometimes, the terms "freedom" and "democracy" are used interchangeably. But democracy, as a set of practices and principles about freedom, rather protects freedom and equality.  
     Measuring democracy also has been contested that there still have ongoing debates on this subject. The most widely-known measure is from the US-based Freedom House organization, which aggregates scores of 25 indicators up to 100, on a 1 to 7 rating scale, from political rights and civil liberties. For the report of 2016, Freedom House evaluated the state of freedom in 195 countries and 15 territories during calendar year 2015, and 125 countries are classified as electoral democracies. The Freedom House's electoral democracy measure is a thin or minimalist whose measure of democracy may differ from thick or wider measures which include various aspects of social and political culture in democratic societies. Another organization where also provides measures of democracy and regime types based on minimalist definitions is The Polity Project. Specifically focused on the institutionalized authority patterns of the state regime, The Polity Project "examines concomitant qualities of democratic and autocratic authority in governing institutions [.to] envision a spectrum of governing authority that spans from fully institutionalized autocracies through mixed, or incoherent, authority regimes to fully institutionalized democracies (Center for Systemic Peace). For our current research, this Polity data from the Polity Project will be used. The detailed information with regard to the "Polity Score" will be discussed at the "Data, Variables, and Methods" section. 
     Following the introduction, this research paper will proceed by presenting previous studies to provide background researches about democratization and its history in Africa to set up the baseline and link to our research question and hypotheses at the following section. After that, details about data, variable selections, and overall statistical methods will be addressed along with explanations how the authors collected, cleaned, and utilized each variable. Based on the cleaned data, the authors will provide descriptive analyses as a preliminary stage, and then the multivariate analyses conducted based on the various statistical models will be interpreted by concentrating on the effects of five socio-economic variables on democratization in three selected African countries. 

# Prior studies and background about Democratization
      Democratization, according to "An Agenda for Democratization" by Boutros Boutros-Ghali, the formal Secretary General of the United Nations, "is a process which leads to a more open, more participatory, less authoritarian society" (1) within a sovereign state. Between the years 1974 and 1990, the world experienced "Third Wave" of democratization which had changed 30 countries from authoritarian regimes to democracies. The penetration and success of democratization in this time period displayed different examples and outcomes, and this raised the questions for the causes of change between political regimes and how the process of democratization sustained in non-democratic societies. 
      From his On Democracy, Robert Dahl suggests three conditions that are essential to attain democratic institutions through democratization and they are: control over military and police forces by elected officials; the absence of foreign intervention opposing democratization; democratic beliefs and political culture; [possibly] modern market economy and economic growth; and the absence of cultural pluralism. Similarly, Samuel Huntington, from his book The Third Wave, argues that democracy could be realized through one of three main types of democratization process and they are: transformation led primarily by the ruling elites; replacement where the opposition plays the main role; and transplacement, a product of active engagement of both sides. With looking into these arguments, it can be claimed that there is no single precondition or factor which perfectly lead the country to be democratized. It has to be a combination of features which should be uniquely and country specifically taken since the strength of each factor may bring different impacts country by country.   
      Africa has also been the wave of political transitions from various types of dictatorships to more open political systems by accepting democracy as a political alternative. However, to understand the democracy and the democratization in Africa, its background of "the relatively recent end to colonial rule [.], post-colonial utopian African socialism, [.] and economic regression in the 1980s" (Samarasinghe, 1994) has to be taken into account. In spite of all these negative factors, many of African countries has been undergoing the democratization based on demand-driven activities, even though the degree of commitment and speed of transition vary significantly. 
  

# Research Question
test


# Data Availability

1) Measurement of Democratization
  1-1) Polity 4
 ?? ??http://www.systemicpeace.org/inscr/p4manualv2015.pdf
  1-2) 

2) Gross National Income Level

3) primary enrollment

4) Income Inequality

5) Gender Inequality in labor force

6) mortality rate under 5

# Democratization

```{r,echo=FALSE}

#we imported data from world bank and UN
dfsa <- read.csv("SA.csv", header = TRUE, sep = ",")
dfbo <- read.csv("BTW.csv", header = TRUE, sep = ",")
dfkn <- read.csv("KNY.csv", header = TRUE, sep = ",")
dfpanel <- read.csv("panel.csv", header = TRUE, sep = ",")
ggplot(dfsa, aes(year, demo)) + geom_line()
ggplot(dfbo, aes(year, demo)) + geom_line()
ggplot(dfkn, aes(year, demo)) + geom_line()


```

# Explanetory variables
|variable name|detail|source|
|:-----|:-----|:------|
|gdppc|Gross National Production Per Capita|World Bank|
|pe|Primary enrollment in education|United Nations|
|mr|Infant Mortaliry Rate|United Nations|
|gi|gender inequality in labor market|United Nations|

# OLS results
```{r star, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}
library(stargazer, quietly = TRUE)

L1 <- lm(demo ~ log(gdppc) + log(pe) + log(mr) + log(gi), data=dfsa)
L2 <- lm(demo ~ log(gdppc) + log(pe) + log(mr) + log(gi), data=dfbo)
L3 <- lm(demo ~ log(gdppc) + log(pe) + log(mr) + log(gi), data=dfkn)

stargazer(L1, L2, L3,digits = 2,column.labels=c("SouthAfrica","Botswana","Kenya"),title = "Regression results for each country", type = 'latex')
```

# Pooled OLS

```{r, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE}
#Then, we conducted pooled OLS. In this model, we included all three countries and run the regression.
L4 <- lm(demo ~ log(gdppc) + log(pe) + log(mr) + log(gi), data=dfpanel)
stargazer(L4,digits = 2,title = "Pooled OLS", type = 'latex')
```

# residual vs fitted value

```{r, echo=FALSE}
plot(fitted(L4), residuals(L4), xlab = "Fitted Values", ylab = "Residuals")
abline(h=0, lty=2)
lines(smooth.spline(fitted(L4), residuals(L4)))
#apparently, this regression suffered from panel specific effects. Because, the 

```

# Breusch-Pagan test
```{r, echo=FALSE}
#apparently, this regression suffered from panel specific effects. Because, the 
bptest(L4)
```


# Fixed-Effect model

```{r, results = 'asis', warning=FALSE, message=FALSE,echo = FALSE}
#the previous regression suffered from panel specific effects, which cause to biased and inefficient estimator. Thus, we use Fixed-effect model to eliminate panel specific effects.

pooled <- plm(demo ~ log(gdppc) + log(pe) + log(mr) + log(gi), data=dfpanel, model="pooling")
fixed <- plm(demo ~ log(gdppc) + log(pe) + log(mr) + log(gi), data=dfpanel, index=c("country", "year"), model="within")

stargazer(pooled, fixed,digits = 3,column.labels=c("PooledOLS","FixedOLS"),title = "pooled OLS and fixed effects OLS", type = 'latex')

#plot(dfpanel$gdppc, res, ylab="Residuals", xlab="GDP per capita", main="country specific effects") 
```
constants:
```{r}
fixef(fixed)
```
# Do panel specific effects exist?

```{r}
pFtest(fixed,pooled)
```

# Breush-Pagan test
```{r, echo=FALSE}
plmtest(fixed,effect="individual",type="bp")
```
we cannot reject the null hypothesis. (residuals doesn't correlated with independent variables)

# Random-Effects OLS
```{r, results = 'asis', warning=FALSE, message=FALSE,echo = FALSE}

random <- plm(demo ~ log(gdppc) + log(pe) + log(mr) + log(gi), data=dfpanel, index=c("country", "year"), model="random")

stargazer(pooled, fixed, random,digits=3 ,title = "random effects OLS", type = 'latex')

```

#Hausman test
```{r, results = 'asis', warning=FALSE, message=FALSE,echo = FALSE}

phtest(fixed, random)
```

# results

```{r, results = 'asis', warning=FALSE, message=FALSE, echo=FALSE,figure_width =5}
options(width = 10)
stargazer(pooled,fixed,random,digits=3, title = "Regression results", type = 'latex')
summary(fixed,type = 'latex')
```

$$ democratization = -1.496 log(gdppc)+15.793log(pe)-12.087log(mr)-28.940log(gi)+ \alpha_i $$

where a_i represents panel specific effects

# Conclusion