---
title: "Socio-economic factors for Democracy <South Africa, Botwana, and Kenya compared>                                                                                                                                                                                                               Takuma Ahndo & Bomi Kim"
output: revealjs::revealjs_presentation
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(stargazer)
library(bibtex)

```

## Democracy? Democratization?  

## Background Research 1 - Factors for Democratization 

## Background Research 2 - Democratization in Africa

## Research Question and Hypotheses


## Data gathering and cleaning

## Descriptive statistics

## Democratization in South Africa
```{r,echo=FALSE}
dfsa <- read.csv("SA.csv", header = TRUE, sep = ",")
dfbo <- read.csv("BTW.csv", header = TRUE, sep = ",")
dfkn <- read.csv("KNY.csv", header = TRUE, sep = ",")
ggplot(dfsa, aes(year, demo)) + geom_line()

L1 <- lm(demo ~ gdppc + pe + mr + gi, data=dfsa)
L2 <- lm(demo ~ gdppc + pe + mr + gi, data=dfbo)
L3 <- lm(demo ~ gdppc + pe + mr + gi, data=dfkn)

```

## Explanetory variables
|variable name|detail|source|
|:-----|:-----|:------|
|gdppc|Gross National Production Per Capita|World Bank|
|pe|Primary enrollment in education|United Nations|
|mr|Infant Mortaliry Rate|United Nations|
|gi|gender inequality in labor market|United Nations|

## Summary of SouthAfrica
```{r}
summary(L1)
```

## Regression Analysis for Botswana
```{r}
summary(L2)

```

## Regression Analysis for Kenya
```{r}
summary(L3)

```
## Conclusion



#References

