---
title: "Socio-economic factors for Democracy < South Africa, Botwana, and Kenya compared >  "                           
author: Takuma Andoh & Bomi Kim
output: revealjs::revealjs_presentation
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(stargazer)
library(bibtex)

```

## Democracy? Democratization?  

## Background Research 1 - Factors for Democratization 

## Background Research 2 - Democratization in Africa

## Research Question and Hypotheses


## Analysis Methodology 1 - Collecting and Gathering Data


## Analysis Methodology 2 - Descriptive statistics

## Democratization in South Africa
```{r,echo=FALSE}

#we imported data from world bank and UN
dfsa <- read.csv("SA.csv", header = TRUE, sep = ",")
dfbo <- read.csv("BTW.csv", header = TRUE, sep = ",")
dfkn <- read.csv("KNY.csv", header = TRUE, sep = ",")
dfpanel <- read.csv("panel.csv", header = TRUE, sep = ",")


ggplot(dfsa, aes(year, demo)) + geom_line()

L1 <- lm(demo ~ gdppc + pe + mr + gi, data=dfsa)
L2 <- lm(demo ~ gdppc + pe + mr + gi, data=dfbo)
L3 <- lm(demo ~ gdppc + pe + mr + gi, data=dfkn)

ggplot(dfsa, aes(year, demo)) + geom_line()

L1 <- lm(demo ~ gdppc + pe + mr + gi, data=dfsa)
L2 <- lm(demo ~ gdppc + pe + mr + gi, data=dfbo)
L3 <- lm(demo ~ gdppc + pe + mr + gi, data=dfkn)

```

## Explanetory variables
|variable name|detail|source|
|:-----|:-----|:------|
|gdppc|Gross National Production Per Capita|World Bank|
|pe|Primary enrollment in education|United Nations|
|mr|Infant Mortaliry Rate|United Nations|
|gi|gender inequality in labor market|United Nations|

<<<<<<< HEAD
## Analysis Methodology 3 - Regression Analysis
=======
## Regression Analysis for SouthAfrica
>>>>>>> origin/master
```{r}
summary(L1)
```

<<<<<<< HEAD

## Conclusion

## Reference
=======
## Regression Analysis for Botswana
```{r}
summary(L2)
```

## Regression Analysis for Kenya
```{r}
summary(L3)
```

## Pooled OLS
```{r}
#Then, we conducted pooled OLS. In this model, we included all three countries and run the regression.

L4 <- lm(demo ~ gdppc + pe + mr + gi, data=dfpanel)
summary(L4)
```

## Country specific effects

```{r}
plot(fitted(L4), residuals(L4), xlab = "Fitted Values", ylab = "Residuals")
abline(h=0, lty=2)
lines(smooth.spline(fitted(L4), residuals(L4)))

#apparently, this regression suffered from panel specific effects.
```

## Fixed-Effect model

```{r, echo = FALSE}
#the previous regression suffered from panel specific effects, which cause to biased and inefficient estimator. Thus, we use Fixed-effect model to eliminate panel specific effects.


#plot(dfpanel$gdppc, res, ylab="Residuals", xlab="GDP per capita", main="country specific effects") 
```

## Conclusion



#References

>>>>>>> origin/master




















<p><img src="data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAABgAAAASACAMAAAAOH3aoAAAAllBMVEUAAAAAADoAAGYAOpAAZrYzMzM6AAA6ZmY6kNtNTU1NTW5NTY5NbqtNjshmAABmZmZmtrZmtv9uTU1uq+SOTU2Oq6uOyMiOyP+QOgCQkLaQ2/+rbk2rjk2r5P+2ZgC2kDq2///Ijk3I///bkDrb///kq27k5Kvk///r6+v/tmb/yI7/25D/5Kv//7b//8j//9v//+T///9dzdohAAAACXBIWXMAAB2HAAAdhwGP5fFlAAAgAElEQVR4nO2d63pbx4FlmY7tCWzLM6bVLU/U1jStZtumGVl8/5cb4kIKIG5V5xS4a9de60ck0omD9W2qFi4H0NUDAABEcqW+AQAAoIEAAACEQgAAAEIhAAAAoRAAAIBQCAAAQCgEAAAgFAIAABAKAQAACIUAAACEQgAAAEIhAAAAoRAAAIBQmgfg9...(line truncated)...


































<p>&lt;&lt;&lt;&lt;&lt;&lt;&lt; HEAD ## Analysis Methodology 3 - Regression Analysis ======= ## Regression Analysis for SouthAfrica &gt;&gt;&gt;&gt;&gt;&gt;&gt; origin/master</p>
<div class="sourceCode"><pre class="sourceCode r"><code class="sourceCode r"><span class="kw">summary</span>(L1)</code></pre></div>

## Call:
## lm(formula = demo ~ gdppc + pe + mr + gi, data = dfsa)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -1.53909 -0.16387  0.03986  0.18542  0.82555 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(&gt;|t|)  
## (Intercept)  6.331e+01  2.563e+01   2.470   0.0295 *
## gdppc       -7.301e-04  3.935e-04  -1.855   0.0883 .
## pe           5.341e-08  4.925e-07   0.108   0.9154  
## mr          -1.540e-01  1.040e-01  -1.480   0.1646  
## gi          -6.803e+01  2.242e+01  -3.034   0.0104 *
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.6706 on 12 degrees of freedom
##   (38 observations deleted due to missingness)
## Multiple R-squared:  0.9372, Adjusted R-squared:  0.9162 
## F-statistic: 44.74 on 4 and 12 DF,  p-value: 4.076e-07</code></pre>
<p>&lt;&lt;&lt;&lt;&lt;&lt;&lt; HEAD</p>




<section><section id="reference" class="titleslide slide level1"><h1>## Reference</h1></section><section id="regression-analysis-for-botswana" class="slide level2">
<h1>Regression Analysis for Botswana</h1>
<div class="sourceCode"><pre class="sourceCode r"><code class="sourceCode r"><span class="kw">summary</span>(L2)</code></pre></div>
<pre><code>## 
## Call:
## lm(formula = demo ~ gdppc + pe + mr + gi, data = dfbo)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -0.42502 -0.20434 -0.05208  0.16324  0.54724 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(&gt;|t|)   
## (Intercept) -3.765e+00  3.985e+00  -0.945  0.35871   
## gdppc        4.283e-04  1.427e-04   3.001  0.00847 **
## pe           3.789e-06  5.207e-06   0.728  0.47742   
## mr           3.059e-02  1.102e-02   2.776  0.01350 * 
## gi           9.267e+00  4.715e+00   1.965  0.06698 . 
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.2878 on 16 degrees of freedom
##   (34 observations deleted due to missingness)
## Multiple R-squared:  0.8471, Adjusted R-squared:  0.8089 
## F-statistic: 22.16 on 4 and 16 DF,  p-value: 2.325e-06</code></pre>
</section><section id="regression-analysis-for-kenya" class="slide level2">
<h1>Regression Analysis for Kenya</h1>
<div class="sourceCode"><pre class="sourceCode r"><code class="sourceCode r"><span class="kw">summary</span>(L3)</code></pre></div>
<pre><code>## 
## Call:
## lm(formula = demo ~ gdppc + pe + mr + gi, data = dfkn)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -4.3910 -0.5731  0.1546  0.6748  3.9389 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(&gt;|t|)   
## (Intercept)  1.696e+02  5.029e+01   3.372  0.00456 **
## gdppc       -1.657e-02  9.355e-03  -1.771  0.09836 . 
## pe           2.402e-06  1.732e-06   1.387  0.18707   
## mr          -3.389e-01  9.479e-02  -3.575  0.00304 **
## gi          -1.768e+02  4.471e+01  -3.955  0.00144 **
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 1.905 on 14 degrees of freedom
##   (40 observations deleted due to missingness)
## Multiple R-squared:  0.9267, Adjusted R-squared:  0.9058 
## F-statistic: 44.27 on 4 and 14 DF,  p-value: 8.49e-08</code></pre>
</section><section id="pooled-ols" class="slide level2">
<h1>Pooled OLS</h1>
<div class="sourceCode"><pre class="sourceCode r"><code class="sourceCode r"><span class="co">#Then, we conducted pooled OLS. In this model, we
included all three countries and run the regression.</span>

L4 &lt;-<span class="st"> </span><span class="kw">lm</span>(demo ~<span class="st"> </span>gdppc +<span class="st"> </span>pe +<span class="st"> </span>mr +<span class="st"> </span>gi, <span class="dt">data=</span>dfpanel)
<span class="kw">summary</span>(L4)</code></pre></div>
<pre><code>## 
## Call:
## lm(formula = demo ~ gdppc + pe + mr + gi, data = dfpanel)
## 
## Residuals:
##     Min      1Q  Median      3Q     Max 
## -3.9141 -1.6893 -0.6646  0.6433  8.1385 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(&gt;|t|)    
## (Intercept)  6.918e+01  1.130e+01   6.124 1.23e-07 ***
## gdppc       -1.246e-03  6.213e-04  -2.006  0.05010 .  
## pe          -5.569e-07  1.948e-07  -2.859  0.00610 ** 
## mr          -1.036e-01  3.078e-02  -3.366  0.00144 ** 
## gi          -7.181e+01  1.363e+01  -5.270 2.66e-06 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 3.007 on 52 degrees of freedom
##   (108 observations deleted due to missingness)
## Multiple R-squared:  0.7434, Adjusted R-squared:  0.7237 
## F-statistic: 37.66 on 4 and 52 DF,  p-value: 8.882e-15</code></pre>
</section><section id="country-specific-effects" class="slide level2">
<h1>Country specific effects</h1>
<div class="sourceCode"><pre class="sourceCode r"><code class="sourceCode r"><span class="kw">plot</span>(<span class="kw">fitted</span>(L4), <span class="kw">residuals</span>(L4), <span class="dt">xlab =</span> <span class="st">&quot;Fitted Values&quot;</span>, <span class="dt">ylab =</span> <span class="st">&quot;Residuals&quot;</span>)
<span class="kw">abline</span>(<span class="dt">h=</span><span class="dv">0</span>, <span class="dt">lty=</span><span class="dv">2</span>)
<span class="kw">lines</span>(<span class="kw">smooth.spline</span>(<span class="kw">fitted</span>(L4), <span class="kw">residuals</span>(L4)))</code></pre></div>
<p><img src="data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAABgAAAASACAMAAAAOH3aoAAAAaVBMVEUAAAAAADoAAGYAOpAAZmYAZpAAZrY6AAA6OgA6Ojo6kJA6kLY6kNtmAABmAGZmtrZmtv+QOgCQkGaQkNuQ29uQ2/+2ZgC2Zma2/7a2///bkDrbtmbb/7bb////tmb/25D//7b//9v////PVlcUAAAACXBIWXMAAB2HAAAdhwGP5fFlAAAgAElEQVR4nO3dC3fbxhaYUSZp7T7s1mkbtVEb2dH//5EV3+BTeAyAOXP2XuuuSL4SRA7A+UgABDfvAKS0WfsGALAOAQBISgAAkhIAgKQEACApAQBISgAAkhIAgKQEACApAQBISgAAkhIAgKQEACApAQBISgAAkhIAgKQEACApAQBISgAAkhIAgKQEACApAQBISgAAkhIAgKQEACApAQBISgAAkhIAgKQEACApAQBISgAAkhIAgKQEA...(line truncated)...
<div class="sourceCode"><pre class="sourceCode r"><code class="sourceCode r"><span class="co">#apparently, this regression suffered from panel specific effects.</span></code></pre></div>
</section><section id="fixed-effect-model" class="slide level2">
<h1>Fixed-Effect model</h1>
</section><section id="conclusion-1" class="slide level2">
<h1>Conclusion</h1>
</section></section>
<section><section id="references" class="titleslide slide level1"><h1>References</h1></section></section>
