---
title: "Socio-economic factors for Democracy <South Africa, Botwana, and Kenya compared>                                                                                                                                                                                                               Takuma Ahndo & Bomi Kim"
output: revealjs::revealjs_presentation
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(stargazer)
library(bibtex)

```

## Democracy? Democratization?  

## Background Research 1 - Factors for Democratization 

## Background Research 2 - Democratization in Africa

## Research Question and Hypotheses


## Data gathering and cleaning

## Descriptive statistics

## Democratization in South Africa
```{r,echo=FALSE}

#we imported data from world bank and UN
dfsa <- read.csv("SA.csv", header = TRUE, sep = ",")
dfbo <- read.csv("BTW.csv", header = TRUE, sep = ",")
dfkn <- read.csv("KNY.csv", header = TRUE, sep = ",")
dfpanel <- read.csv("panel.csv", header = TRUE, sep = ",")


ggplot(dfsa, aes(year, demo)) + geom_line()

L1 <- lm(demo ~ gdppc + pe + mr + gi, data=dfsa)
L2 <- lm(demo ~ gdppc + pe + mr + gi, data=dfbo)
L3 <- lm(demo ~ gdppc + pe + mr + gi, data=dfkn)

```

## Explanetory variables
|variable name|detail|source|
|:-----|:-----|:------|
|gdppc|Gross National Production Per Capita|World Bank|
|pe|Primary enrollment in education|United Nations|
|mr|Infant Mortaliry Rate|United Nations|
|gi|gender inequality in labor market|United Nations|

## Regression Analysis for SouthAfrica
```{r}
summary(L1)
```

## Regression Analysis for Botswana
```{r}
summary(L2)
```

## Regression Analysis for Kenya
```{r}
summary(L3)
```

## Pooled OLS
```{r}
#Then, we conducted pooled OLS. In this model, we included all three countries and run the regression.

L4 <- lm(demo ~ gdppc + pe + mr + gi, data=dfpanel)
summary(L4)
```

## Country specific effects

```{r}
plot(fitted(L4), residuals(L4), xlab = "Fitted Values", ylab = "Residuals")
abline(h=0, lty=2)
lines(smooth.spline(fitted(L4), residuals(L4)))

#apparently, this regression suffered from panel specific effects.
```

## Fixed-Effect model

```{r, echo = FALSE}
#the previous regression suffered from panel specific effects, which cause to biased and inefficient estimator. Thus, we use Fixed-effect model to eliminate panel specific effects.


#plot(dfpanel$gdppc, res, ylab="Residuals", xlab="GDP per capita", main="country specific effects") 
```

## Conclusion



#References

