---
title: "Presentation"
output: revealjs::revealjs_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(stargazer)
```

## Socio-economic factors for Democracy <South Africa, Botwana, and Kenya compared> 

Takuma Ahndo & Bomi Kim
        Fall 2016 

## Democracy? Democratization?  

## Background Research 1 - Factors for Democratization 

## Background Research 2 - Democratization in Africa

## Research Question and Hypotheses


## Data gathering and cleaning

## Descriptive statistics

## Democratization in South Africa
```{r,echo=FALSE}
df <- read.csv("SA.csv", header = TRUE, sep = ",")
ggplot(df, aes(year, demo)) + geom_line()
```

## Explanetory variables
|variable name|detail|source|
|:-----|:-----|:------|
|gdppc|Gross National Production Per Capita|World Bank|
|pe|Primary enrollment in education|United Nations|
|mr|Infant Mortaliry Rate|United Nations|
|gi|gender inequality in labor market|United Nations|

## Regression Analysis
```{r}
fit3 <- lm(demo ~ gdppc + pe+mr+gi, data=df)
stargazer(fit3,type="text")
```

## 

## Conclusion
